---
title: "Multivarible Analysis"
author: "Group 4: Aishe Memetova, Haotian Bu, Jon White, Vitalii Zakhozhyi"
date: "March 9, 2019"
output: 
  html_document:
    toc: true
    toc_depth: 1
---

______

Using the Seattle Sound Transit Data of October-November 2015, we will try to explore factors which may affect people's probability of using cash for the trip.


#Intro: Loading Data and Describing the Data

Loading neccessary packages, calling dataset and preparing it for analysis
```{r warning = FALSE}
library(ggplot2)
library(scales)
mydata1=read.csv("https://raw.githubusercontent.com/karllovepolicy/Data_Visual_Final_Tutorial/master/FinalData.csv")
```

```{r echo=FALSE, include=FALSE, warning = FALSE}
str(mydata1)
```


#Multi-variable plot

In this section, we will explore the realationship among the probability of using cash for this trip, whether people have vehicles, the purpose of this trip, and frequency of using public transportation. 

First, we need to make our dataset ready for regression. Following code select dependent variables and explanatory variables, then set the right format for each variable.

```{r multi transform binamiry variable_cash, message=FALSE, warning = FALSE}
#change names of explanatory variables
library(dplyr)
mydata1 <- rename(mydata1,'number_of_trips'='oneway_all_clean', 'vehicle'='hhveh_clean')

# transform explanatory varibles from factor to numeric, add "-1" in command because Factors in R are represented internally as integers and the first factor is 1.
mydata1$vehicle <- as.numeric(mydata1$vehicle)-1
mydata1$number_of_trips <- as.numeric(mydata1$number_of_trips)-1

# subset, select needed variables:
mydata2 <- mydata1 %>% mutate(cashlove = ifelse(faretype=="Cash", 1, 0), vehicle = ifelse(vehicle>0, 1, 0)) %>% select(cashlove, purpose, number_of_trips, vehicle) %>% filter(purpose != "#NULL!")
# set the correct type for explanatory variables
mydata2$cashlove <- factor(mydata2$cashlove)
mydata2$vehicle <- factor(mydata2$vehicle)
mydata2$purpose <- factor(mydata2$purpose) #reset the level of variable "purpose"
mydata2$number_of_trips <- as.numeric(mydata2$number_of_trips)
head(mydata2,5)
```

Here we use logit model since our dependent variable is a binary variable. Our reference group is people whose purpose of this trip is home-based non-work trip and don't have vehicles in home.
```{r logit regression, message=FALSE, warning=FALSE}
library(dotwhisker)
library(broom)
#just use purpose as explantary variable
mylogit1 <- glm(cashlove ~ purpose, data = mydata2, family = "binomial")
model1 = tidy(mylogit1) %>%   # we save the result as a tidy object and...
    mutate(model = "Model 1")
#use purpose and number of trips
mylogit2 <- glm(cashlove ~ purpose + number_of_trips, data = mydata2, family = "binomial")
model2 = tidy(mylogit2) %>%   # we save the result as a tidy object and...
    mutate(model = "Model 2")
#use all three explanatory varibales
mylogit3 <- glm(cashlove ~ purpose + number_of_trips + vehicle, data = mydata2, family = "binomial")
model3 = tidy(mylogit3) %>%   # we save the result as a tidy object and...
    mutate(model = "Model 3")

summary(mylogit3)
```


Comparison among three models:
```{r compare different models}
allModels=rbind(model1, model2, model3)

dwplot(allModels, vline = geom_vline(xintercept = 0, colour = "grey60", linetype = 2)) + 
  xlab("Coefficient Estimate") + 
  ylab("") +
  ggtitle("Predicting probability of using cash") +
  theme_bw()
```

From plot above, we can see that compared with reference group, people whose purpose is home based work trip have less probability using cash, while people whose purpose is non-home based trip is more likely to use cash.
The more trips people have in each month, the less likelihood people would use cash. However, the difference on cash usage between people who have vehicles in home and people who don't is non-significant. 

